Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-12 surfaces-write-up/surface-write-up.tex
hirzebruch macro, change a, b -> c, d in 1pt

\DeclareMathOperator\di{Div}
\newcommand\bida{a}
\newcommand\bidb{b}

@@

\todo{macros: hirzebruch, degrees}

@@

\begin{thm}
\label{thm:proj-one-point}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \alpha H$, with $\alpha = \frac{a}{b} \in \bq$
and $H = V(x_k)$ a hyperplane of $\bp^m$, be a divisor on $\bp^m$.
Let
\[
	\left\lfloor \frac{a}{b} \right\rfloor = \frac{c_0}{d_0} <
	\frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{a}{b}
\]

@@

Now we show that no further generators are necessary. As done in
the proof of the $\bp^1$ case by O'Dorney \cite[Theorem 6]
{dorney:canonical}, each lattice point $(d, c) \in \bz_{\geq 0}^2$
such that $c \leq d \alpha$ is a non-negative linear combination of
$\vec{x}_i$ and $\vec{x}_{i + 1}$ where $\vec{x}_i := (d_i, c_i) \in
\bz^2$. Consequently, any element $\frac{u^{d} x^{\vec{v}}}
{x_k^{c}} \in R_D$ is expressible as
\begin{align*}
	\frac{u^{d} x^{\vec{v}}} {x_k^{c}} = \frac{u^{d_i} x^{\vec{w}}}
	{x_k^{c_i}} \frac{u^{d_{i + 1}} x^{\vec{z}}} {x_k^{c_{i + 1}}}
\end{align*}

>>>>

\DeclareMathOperator\di{Div}
\newcommand\bida{a}
\newcommand\bidb{b}
\newcommand\hirz{\mathbb{F}}

@@

\todo{macros: degrees}

@@

\begin{thm}
\label{thm:proj-one-point}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \alpha H$, with $\alpha = \frac{a}{b} \in \bq$
and $H = V(x_k)$ a hyperplane of $\bp^m$, be a divisor on $\bp^m$.
Let
\[
	\left\lfloor \frac{c}{d} \right\rfloor = \frac{c_0}{d_0} <
	\frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{c}{d}
\]

@@

Now we show that no further generators are necessary. As done in
the proof of the $\bp^1$ case by O'Dorney \cite[Theorem 6]
{dorney:canonical}, each lattice point $(d', c') \in \bz_{\geq 0}^2$
such that $c' \leq d' \alpha$ is a non-negative linear combination of
$\vec{x}_i$ and $\vec{x}_{i + 1}$ where $\vec{x}_i := (d_i, c_i) \in
\bz^2$. Consequently, any element $\frac{u^{d'} x^{\vec{v}}}
{x_k^{c'}} \in R_D$ is expressible as
\begin{align*}
	\frac{u^{d'} x^{\vec{v}}} {x_k^{c'}} = \frac{u^{d_i} x^{\vec{w}}}
	{x_k^{c_i}} \frac{u^{d_{i + 1}} x^{\vec{z}}} {x_k^{c_{i + 1}}}
\end{align*}
